---
title: "Species_response"
author: "Ashal Shrestha, ID 800920396"
date: "2025-11-21"
output: html_document
---
#Load libraries
```{r}
library(dplyr)
library(stringr)
library(ggplot2)
library(lme4)

```




```{r}
setwd("C:/My_analysis")

leaf <- read.csv("leaf_traits.csv",header = TRUE)

root <- read.csv("root_traits.csv", header = TRUE)
type <- read.csv("spp_list.csv", header = TRUE)

#abundance_npk <- read.csv("C:/My_analysis/Nut_net_data_mgmt/generated_data/npk_control_differences.csv", header = TRUE)

#abundance_npk_fence <- read.csv("C:/My_analysis/Nut_net_data_mgmt/generated_data/npk_fence_control_differences.csv", header = TRUE)

#abundance_fence <- read.csv("C:/My_analysis/Nut_net_data_mgmt/generated_data/fence_control_differences.csv", header = TRUE)

abundance_all <- read.csv("C:/My_analysis/Nut_net_data_mgmt/generated_data/all_trt_abundance_differences_upto2024.csv", header = TRUE)
```

#Get the required column only
```{r}
#Get the required column only
type<- type[c("Taxon" ,"species")]
#Remove duplicate Taxon
type_clean <- type[!duplicated(type$Taxon),]


#############Rename the taxon that are duplicated in the species lists.
#From ELLIOT, south Cali
type_clean$Taxon[type_clean$Taxon == "LOLIUM MULTIFLORUM" & type_clean$site_id == "elliot.us"] <- "LOLIUM PERENNE"
type_clean$species[type_clean$species == "" & type_clean$site_id == "elliot.us"] <- "FEPE"

#From SGS, colorado
type_clean$Taxon[type_clean$Taxon == "ELYMUS CANADENSIS"] <- "ELYMUS ELYMOIDES"
type_clean$species[type_clean$species == "ELCA"] <- "ELEL"


#############From KONZ, Kansas
#type_clean$Taxon[type_clean$Taxon == "CHENOPODIUM BERLANDIERI"] <- "CHENOPODIUM ALBUM"
#type_clean$species[type_clean$species == "CHBE"] <- "CHAL"

#type_clean$Taxon[type_clean$Taxon == "MUHLENBERGIA RACEMOSA"] <- "MUHLENBERGIA CUSPIDATA"
#type_clean$species[type_clean$species == "MURA"] <- "MUCU"

#type_clean$Taxon[type_clean$Taxon == "SPOROBOLUS CRYPTANDRUS"] <- "SPOROBOLUS COMPOSITUS"
#type_clean$species[type_clean$species == "SPCR"] <- "SPCOMP"

#From Cdcr, Minnesota
type_clean$Taxon[type_clean$Taxon == "LYCHNIS LATIFOLIA SSP. ALBA" & type_clean$site_id == "cdcr.us"] <- "SILENE LATIFOLIA"
type_clean$species[type_clean$species == "LYLA" & type_clean$site_id == "crdr.us"] <- "SILA"
type_clean$species[type_clean$species == "" & type_clean$Taxon == "LYCHNIS LATIFOLIA SSP. ALBA"] <- "SILA"


type_clean <- type_clean[!duplicated(type_clean$species),]
type_clean <- type[!is.na(type$species) & type$species != "", ]     #Remove NAs
type_data <- type_clean[!duplicated(type_clean$species), ]       #handle duplicates
```



#Refine_leaves_dataframe
```{r}
leaves<- leaf
#leaves$find_outlier <- leaves$Sample_ID          #We need a code later to see the samples with outliers

leaves <- leaves %>%
  mutate(
    SITE_ID = case_when(
      grepl("hopl", SITE_ID, ignore.case = TRUE) ~ "Hopl.us",
      grepl("sier", SITE_ID, ignore.case = TRUE) ~ "Sier.us",
      grepl("elliot", SITE_ID, ignore.case = TRUE) ~ "Elliot.us",
      grepl("sevi", SITE_ID, ignore.case = TRUE) ~ "Sevi.us",
      grepl("konz", SITE_ID, ignore.case = TRUE) ~ "Konz.us",
      grepl("sgs", SITE_ID, ignore.case = TRUE) ~ "Sgs.us",
      grepl("spin", SITE_ID, ignore.case = TRUE) ~ "Spin.us",
      grepl("kbs", SITE_ID, ignore.case = TRUE) ~ "Kbs.us",
      grepl("crdr", SITE_ID, ignore.case = TRUE) ~ "Cdcr.us",
      grepl("bnch", SITE_ID, ignore.case = TRUE) ~ "Bnch.us",
      grepl("nmsu", SITE_ID, ignore.case = TRUE) ~ "Spin.us",
      grepl("mcla", SITE_ID, ignore.case = TRUE) ~ "Mcla.us",
      TRUE ~ NA_character_
    )
  )
```

#Get rid of unneccesary columns from df leaves
```{r}
leaves<- leaves[c("SAMPLE_ID", "SITE_ID", "HT_mm", "Leaf_num", "LWS", "LWD", "Leaf_area", "RWF" ,"RWD")]
```


```{r}

#leaves$SITE_ID[leaves$SITE_ID == "Crdr.us"] <- "Cdcr.us"      #Rename one of the site name.
#leaves$SITE_ID[leaves$SITE_ID == "spin.us"] <- "Spin.us"      #Rename one of the site name.
#leaves$sample_name <- paste(leaves$SAMPLE_ID, leaves$SITE_ID, sep = "_")
#leaves$sample_name <- paste(leaves$SAMPLE_ID,tolower(sub("\\.us$", "", leaves$SITE_ID)),  sep = "_")  

names(leaves)[names(leaves) == "SAMPLE_ID"] <- "Sample_ID"   #Rename the column name

leaves <- leaves %>%
  mutate(key = paste(Sample_ID, SITE_ID, sep = "_"))  #making "key" column to merge with other files later

leaves<- leaves[c("key", "HT_mm","Sample_ID", "SITE_ID", "Leaf_num", "LWS", "LWD", "Leaf_area", "RWF" ,"RWD")]    #Remove Sample_ID and SITE_ID to help swift merge later with the roots
```



#Fix the issues with Sample_IDs (Make changes in the column "key") that are duplicated but denotes different species.We have to do this with the root data too. 
```{r}
#BRMA.hopl>BRMAXIMA.hopl
leaves$key <- ifelse(
  grepl("^BRMA_.*_hopl", leaves$key, ignore.case = TRUE),
  sub("^BRMA_", "BRMAXIMA_", leaves$key),
  leaves$key
)

#EUSE.konz>EUSERPENS.kona
leaves$key <- ifelse(
  grepl("^EUSE_.*_Konz", leaves$key, ignore.case = TRUE),
  sub("^EUSE_", "EUSERPENS_", leaves$key),
  leaves$key
)


#EUSE.konz>EUSERPENS.kona
leaves$key <- ifelse(
  grepl("^CGEL_.*_Spin", leaves$key, ignore.case = TRUE),
  sub("^CGEL_", "CEGL_", leaves$key),
  leaves$key
)                                       #it must be CERASTIUM GLOMERATUM

#LECA.cdcr>LECAP.cdcr
leaves$key <- ifelse(
  grepl("^LECA_.*_Cdcr", leaves$key, ignore.case = TRUE),
  sub("^LECA_", "LECAP_", leaves$key),
  leaves$key
)


#MAGR.sevi > MACGRA.sevi (I tried to do MAGR.hopl > MAGRACILIS.hopl first, but I think there is no MAGR or it was not collected from hopl
leaves$key <- ifelse(
  grepl("^MAGR_.*_Sevi", leaves$key, ignore.case = TRUE),
  sub("^MAGR_", "MACGRA_", leaves$key),
  leaves$key
)


#There are 3 SPCO for 3 different species. So, I changed 2 of them.

#SPCO.Sgs > SPCOCCONEA.Sgs
leaves$key <- ifelse(
  grepl("^SPCO_.*_Sgs", leaves$key, ignore.case = TRUE),
  sub("^SPCO_", "SPCOCCINEA_", leaves$key),
  leaves$key
)


#SPCO.Konz > SPCOMP.Konz
leaves$key <- ifelse(
  grepl("^SPCO_.*_Konz", leaves$key, ignore.case = TRUE),
  sub("^SPCO_", "SPCOCCINEA_", leaves$key),
  leaves$key
)


leaves$key <- ifelse(
  grepl("^BODA_.*_Sgs", leaves$key, ignore.case = TRUE),
  sub("^BODA_", "BUDA_", leaves$key),
  leaves$key
)                                #I believe this is BUCHLOE DACTYLOIDES

leaves$key <- ifelse(
  grepl("^POPA_.*_Sgs", leaves$key, ignore.case = TRUE),
  sub("^POPA_", "POPR_", leaves$key),
  leaves$key
)                                            #Poa Pratense?

#TRDU.hopl > TRDUBIUM.hopl
leaves$key <- ifelse(
  grepl("^TRDU_.*_Hopl", leaves$key, ignore.case = TRUE),
  sub("^TRDU_", "TRDUBIUM_", leaves$key),
  leaves$key
)


#SHER.sier > SHAR.Sier    #There is no any spp that can be aligned with SHER in N. Cali 
leaves$key <- ifelse(
  grepl("^SHER_.*_Sier", leaves$key, ignore.case = TRUE),
  sub("^SHER_", "SHAR_", leaves$key),
  leaves$key
)

```


#Change the data after reweighing dry leaf and root samples

```{r}
leaves <- leaves %>%
  mutate(
    LWD = case_when(
      key == "BOER_4_Sevi.us" ~ 0.01468,
      key == "PODO_2_Bnch.us" ~ 0.03285,
      key == "RUHI_3_Cdcr.us" ~ 0.02367,
      key == "POPR_4_Konz.us" ~ 0.04885,
      key == "IRMA_5_Hopl.us" ~ 0.43403,
      key == "ELRE_1_Spin.us" ~ 0.19136,
      key == "GAPA_5_Hopl.us" ~ 0.00039,
      key == "NAPU_1_Hopl.us" ~ 0.18613,
      key == "JUDU_2_Elliot.us" ~ 0.29767,
      key == "CGEL_2_Sier.us" ~ 0.00199,
      key == "AGHE_2_Mcla.us" ~ 0.00145,
      key == "AGHE_6_Mcla.us" ~ 0.00358,
      key == "CLAPER_5_Hopl.us" ~ 0.00570,
      key == "AGHE_1_Mcla.us" ~ 0.00445,
      key == "BRMI_4_Hopl.us" ~ 0.00165,
      key == "VERONICA_5_Kbs.us" ~ 0.00378,
      key == "VERONICA_4_Kbs.us" ~ 0.00230,
      key == "BRMI_3_Hopl.us" ~ 0.00996,
      key == "VERONICA_3_Kbs.us" ~ 0.00659,
      key == "VERONICA_1_Spin.us" ~ 0.00107,
      TRUE ~ LWD  # Keep original value if no match
    ))

leaves <- leaves %>%
  mutate(
    RWD = ifelse(key == "EUAL_2_Sevi.us", 0.00107, RWD)    #Root dry weight
  )

#Change the root dry weight after reweighing (I did this again after team meeting)

leaves <- leaves %>%
  mutate(
    RWD = case_when(
      key == "DAPU_2_Hopl.us" ~ 0.00036,
      key == "ERCI_5_Sier.us" ~ 0.00009,
      key == "VUMI_1_Mcla.us" ~ 0.00005,
      key == "FIGA_1_Hopl.us" ~ 0.00022,
      key == "VISA_1_Spin.us" ~ 0.00095,
      key == "TACA_3_Mcla.us" ~ 0.00035,
      key == "TACA_5_Mcla.us" ~ 0.00008,
      key == "DAPU_4_Hopl.us" ~ 0.00015,
      key == "HECO_3_Mcla.us" ~ 0.00022,
      key == "BRMA_4_Hopl.us" ~ 0.00008,
      key == "EUAL_2_Sevi.us" ~ 0.00010,
      key == "CLAPER_6_Hopl.us" ~ 0.00094,
      key == "BRMAXIMA_4_Hopl.us" ~ 0.00008,
      TRUE ~ RWD
    )
  )

##For root weight fresh
leaves <- leaves %>%
  mutate(
    RWF = case_when(
      key == "DAPU_2_Hopl.us" ~ 0.00144,
     TRUE ~ RWF
    ))

```



#make NAs to the crispy leaves and roots that scale did not detect.
```{r}
#"IRMA_5_Hopl.us"> the dry leaf weight was higher than the saturated, 
#"ELRE_1_Spin.us"> There was not much difference between the dry ans saturated leaf weight, which is very unlikely.
#"PHPR_1,2,5_Kbs.us" are removed because the leaves are crispy when collected.
#"ELRE_3_Kbs.us", "ELRE_1_Kbs.us" are more than 80% crispy leaves when collected
#"POBU_1_Mcla.us", "POBU_2_Mcla.us" these root samples are the one that are not identified
#"FEPE_7_Elliot.us", overly outlier root tissue density. 

keys_to_na_leaf <- c("PHPR_1_Kbs.us","PHPR_2_Kbs.us", "PHPR_5_Kbs.us", "ELRE_3_Kbs.us", "ELRE_1_Kbs.us", "IRMA_5_Hopl.us", "ELRE_1_Spin.us", "LOMU_3_Mcla.us","POBU_1_Mcla.us", "POBU_2_Mcla.us")
leaves$LWS[leaves$key %in% keys_to_na_leaf] <- NA
leaves$LWD[leaves$key %in% keys_to_na_leaf] <- NA


#Remove the dry roots that are not detected by the scale
keys_to_na_roots <- c("LOMU_2_Mcla.us","PLNO_5_Hopl.us", "CAIN_3_Konz.us", "POBU_1_Mcla.us", "FEPE_7_Elliot.us", "BRDI_3_Elliot.us" )
leaves$RWD[leaves$key %in% keys_to_na_roots] <- NA
leaves$RWF[leaves$key %in% keys_to_na_roots] <- NA     #Removed LOMU_3 because dry root weight was 0.00000 that is biasing the results, Removed POBU because there was only root sample and found it not very good sample to keep that can distort our result.

leaves[leaves$key %in% keys_to_na_leaf, c("key", "LWS", "LWD")]
leaves[leaves$key %in% keys_to_na_roots, c("key", "RWF", "RWD")]
```


#Change the leaf area data after re working in ImageJ
```{r}

leaves <- leaves %>%
  mutate(
    Leaf_area = case_when(
      key == "RUCR_5_Spin.us" ~ 154.377,
      key == "PLMA_2_Spin.us" ~ 133.057,
      key == "PLMA_1_Spin.us" ~ 145.220,
      key == "PLMA_3_Spin.us" ~ 108.756,
      key == "RUCR_1_Spin.us" ~ 76.695,
      key == "PLMA_4_Spin.us" ~ 68.792,
      key == "SOHA_5_Spin.us" ~ 99.901,
      key == "SORU_2_Cdcr.us" ~ 91.297,
      key == "SOHA_1_Spin.us" ~ 87.899,
      key == "SOHA_3_Spin.us" ~ 86.030,
      key == "SOHA_4_Spin.us" ~ 90.120,
      key == "SORI_3_Cdcr.us" ~ 75.756,
      key == "ASOV_3_Cdcr.us" ~ 89.647,
      key == "SOHA_2_Spin.us" ~ 78.183,
      key == "CIAR_5_Spin.us" ~ 82.487,
      key == "CIAR_1_Spin.us" ~ 91.58,
      key == "CIAR_2_Spin.us" ~ 71.435,
      key == "SOSP_1_Cdcr.us" ~ 58.645,
      key == "CIAR_4_Spin.us" ~ 62.702,
      TRUE ~ Leaf_area
    )
  )

```


#unit conversion to align with Kattge et al. 2019
```{r}
leaves$Leaf_num<- as.numeric(leaves$Leaf_num)
leaves<-leaves%>% mutate(height_m = HT_mm*0.001) #plant height unit changed from mm to meters.
leaves<-leaves%>% mutate(LWS_mg = LWS*1000)      #leaf weight saturated unit changed from gram to milligram.
leaves<-leaves%>% mutate(LWD_mg = LWD*1000)      #leaf weight dry unit changed from gram to milligram.
leaves<-leaves%>% mutate(RWF_mg = RWF*1000)      #root weight fresh unit changed from gram to milligram.
leaves<-leaves%>% mutate(RWD_mg = RWD*1000)      #root weight dry unit changed from gram to milligram.
leaves<-leaves%>% mutate(Leaf_area_mm2 = Leaf_area*100)        #leaf area unit changed from cm2 to mm2.
#area<-area%>% mutate(Leaf_area_mm2 = Area*100)        #leaf area unit changed from cm2 to mm2.
```

```{r}
write.csv(leaves, "C:/My_analysis/Nut_net_data_mgmt/generated_data/leaves.csv", row.names = FALSE)
```


#Refine_root_dataframe
```{r}
roots<-root

roots <- roots %>%
  mutate(
    Sample_ID = sub("^([^_]+_[^_]+).*", "\\1", File.Name)       
  )                                                        #creat new column named Sample_ID (example: ACAM_1_Elliot.us_ -> ACAM_1 .


roots <- roots %>%
  mutate(
    SITE_ID = case_when(
      grepl("hopl", File.Name, ignore.case = TRUE) ~ "Hopl.us",
      grepl("sier", File.Name, ignore.case = TRUE) ~ "Sier.us",
      grepl("elliot", File.Name, ignore.case = TRUE) ~ "Elliot.us",
      grepl("sevi", File.Name, ignore.case = TRUE) ~ "Sevi.us",
      grepl("konz", File.Name, ignore.case = TRUE) ~ "Konz.us",
      grepl("sgs", File.Name, ignore.case = TRUE) ~ "Sgs.us",
      grepl("spin", File.Name, ignore.case = TRUE) ~ "Spin.us",
      grepl("kbs", File.Name, ignore.case = TRUE) ~ "Kbs.us",
      grepl("crdr", File.Name, ignore.case = TRUE) ~ "Cdcr.us",
      grepl("bnch", File.Name, ignore.case = TRUE) ~ "Bnch.us",
      grepl("nmsu", File.Name, ignore.case = TRUE) ~ "Spin.us",
      grepl("mcla", File.Name, ignore.case = TRUE) ~ "Mcla.us",
      TRUE ~ NA_character_
    )
  )

roots <- roots %>%
  mutate(key = paste(Sample_ID, SITE_ID, sep = "_"))  #making key column to merge with leaves df

unique(roots$SITE_ID) 
```

#Fix the issues with Sample_IDs (Make changes in the column "key") that are duplicated but denotes different species.We have to do this to align with leaves dataframe. 
```{r}
#BRMA.hopl>BRMAXIMA.hopl
roots$key <- ifelse(
  grepl("^BRMA_.*_hopl", roots$key, ignore.case = TRUE),
  sub("^BRMA_", "BRMAXIMA_", roots$key),
  roots$key
)

#EUSE.konz>EUSERPENS.kona
roots$key <- ifelse(
  grepl("^EUSE_.*_Konz", roots$key, ignore.case = TRUE),
  sub("^EUSE_", "EUSERPENS_", roots$key),
  roots$key
)


#LECA.cdcr>LECAP.cdcr
roots$key <- ifelse(
  grepl("^LECA_.*_Cdcr", roots$key, ignore.case = TRUE),
  sub("^LECA_", "LECAP_", roots$key),
  roots$key
)


#MAGR.sevi > MACGRA.sevi (I tried to do MAGR.hopl > MAGRACILIS.hopl first, but I think there is no MAGR or it was not collected from hopl
roots$key <- ifelse(
  grepl("^MAGR_.*_Sevi", roots$key, ignore.case = TRUE),
  sub("^MAGR_", "MACGRA_", roots$key),
  roots$key
)


#There are 3 SPCO for 3 different species. So, I changed 2 of them.

#SPCO.Sgs > SPCOCCONEA.Sgs
roots$key <- ifelse(
  grepl("^SPCO_.*_Sgs", roots$key, ignore.case = TRUE),
  sub("^SPCO_", "SPCOCCINEA_", roots$key),
  roots$key
)


#SPCO.Konz > SPCOMP.Konz
roots$key <- ifelse(
  grepl("^SPCO_.*_Konz", roots$key, ignore.case = TRUE),
  sub("^SPCO_", "SPCOCCINEA_", roots$key),
  roots$key
)


roots$key <- ifelse(
  grepl("^BODA_.*_Sgs", roots$key, ignore.case = TRUE),
  sub("^BODA_", "BUDA_", roots$key),
  roots$key
)                                #I believe this is BUCHLOE DACTYLOIDES


roots$key <- ifelse(
  grepl("^POPA_.*_Sgs", roots$key, ignore.case = TRUE),
  sub("^POPA_", "POPR_", roots$key),
  roots$key
)                                            #Poa Pratense?

#TRDU.hopl > TRDUBIUM.hopl
roots$key <- ifelse(
  grepl("^TRDU_.*_Hopl", roots$key, ignore.case = TRUE),
  sub("^TRDU_", "TRDUBIUM_", roots$key),
  roots$key
)

#MOEFAL.kbs > MEOFAL.kbsl
roots$key <- ifelse(
  grepl("^MOEFAL_.*_kbs", roots$key, ignore.case = TRUE),
  sub("^MOEFAL_", "MEOFAL_", roots$key),
  roots$key
)


#TOAR.kbs > MEOFAL.kbsl
roots$key <- ifelse(
  grepl("^TOAR_.*_Cdcr", roots$key, ignore.case = TRUE),
  sub("^TOAR_", "TORA_", roots$key),
  roots$key
)

#SHER.sier > SHAR.Sier    #There is no any spp that can be aligned with SHER in N. Cali 
roots$key <- ifelse(
  grepl("^SHER_.*_Sier", roots$key, ignore.case = TRUE),
  sub("^SHER_", "SHAR_", roots$key),
  roots$key
)
```


```{r}
roots<- roots[c("Sample_ID","SITE_ID",  "Total.Root.Length.mm", "Network.Area.mm2", "Average.Diameter.mm", "Volume.mm3", "key")]
roots<-roots%>% mutate(Total.Root.Length.cm = Total.Root.Length.mm*0.1)        #root lenght unit changed from mm to cm.
```


```{r}
write.csv(roots, "C:/My_analysis/Nut_net_data_mgmt/generated_data/roots.csv", row.names = FALSE)
```


#Merge leaves and roots dataframe based on column sample_name as a key column. 
```{r}
leaf_root <- left_join(leaves, roots, by = "key")

leaf_root$species <- sub("_.*", "", leaf_root$key)   #To merge later with type df and may be calculate the average values of each spp later
names(leaf_root)[names(leaf_root) == "Sample_ID.x"] <- "Sample_ID"   #Rename the column name
names(leaf_root)[names(leaf_root) == "SITE_ID.x"] <- "SITE_ID"   #Rename the column name
names(leaf_root)
```

#Filter out unneccesary column
```{r}

leaf_root <- leaf_root[, c("Sample_ID", "SITE_ID","LWS", "LWD", "Leaf_num", "RWF", "RWD","LWS_mg", "LWD_mg","RWF_mg", "RWD_mg", "Leaf_area_mm2", "key", "height_m", "Total.Root.Length.mm", "Network.Area.mm2", "Average.Diameter.mm",  "Volume.mm3",  "Total.Root.Length.mm", "species")]                # keep relevant columns 

names(leaf_root)
```

#Join another dataframe named type_data to see the species name.

```{r}
leaf_root <- leaf_root %>% left_join(type_data, by = "species")
unique(leaf_root)

leaf_root <- leaf_root[, c("Sample_ID", "SITE_ID","LWS", "LWD", "Leaf_num", "RWF", "RWD","LWS_mg", "LWD_mg","RWF_mg", "RWD_mg", "Leaf_area_mm2", "key", "height_m", "Total.Root.Length.mm", "Network.Area.mm2", "Average.Diameter.mm",  "Volume.mm3",  "Total.Root.Length.mm", "species", "Taxon")]                # keep relevant columns 
#names(leaf_root)[names(leaf_root) == "Taxon.y"] <- "Taxon"   #Rename the column name
```

#Convert 0's to NAs
```{r}
num_cols <- sapply(leaf_root, is.numeric)

leaf_root[num_cols] <- lapply(
  leaf_root[num_cols],
  function(x) replace(x, x == 0, NA)
)
```


#Leaf traits
```{r}
leaf_root<-leaf_root %>% mutate(SLA_mm2_mg = Leaf_area_mm2/LWD_mg)    #SLA calculation
leaf_root<-leaf_root %>% mutate(LDMC_g_g = LWD_mg/LWS_mg)      #LDMC calculation
leaf_root<-leaf_root %>% mutate(LA_mm2_no = Leaf_area_mm2/Leaf_num)  #Leaf area calculation
leaf_root<-leaf_root %>% mutate(RDMC_g_g = RWD_mg/RWF_mg)      #RDMC calculation
leaf_root<-leaf_root %>% mutate(RTD_gm_mm3 = RWD_mg/Volume.mm3)      #RDMC calculation
leaf_root <-leaf_root %>% mutate(SRL_cm_g = Total.Root.Length.mm/RWD_mg)    #to calculate SRL
leaf_root <-leaf_root %>% mutate(RD_cm = Average.Diameter.mm/10)    #to calculate Root diameter
#summary(leaf_root$SLA_mm2_mg)
```

```{r}
summary(leaf_root[, c("SLA_mm2_mg", "LDMC_g_g", "LA_mm2_no", "RDMC_g_g", "RTD_gm_mm3", "SRL_cm_g", "RD_cm", "height_m")])
```


```{r}
hist(leaf_root$SLA_mm2_mg)
hist(leaf_root$LDMC_g_g)
hist(leaf_root$LA_mm2_no)
hist(leaf_root$RDMC_g_g)
hist(leaf_root$RTD_gm_mm3)
hist(leaf_root$SRL_cm_g)
hist(leaf_root$RD_cm)
hist(leaf_root$height_m)
```



Keep the relevant column in leaf_root.
```{r}
leaf_root <- leaf_root[,c("Sample_ID", "SITE_ID", "key", "Total.Root.Length.mm",  "Network.Area.mm2", "Average.Diameter.mm", "Volume.mm3", "Total.Root.Length.mm", "species", "Taxon", "height_m", "SLA_mm2_mg", "LDMC_g_g", "LA_mm2_no", "RDMC_g_g", "RTD_gm_mm3", "SRL_cm_g", "RD_cm")]
```


#combine above and below ground traits
```{r}
write.csv(leaf_root, "C:/My_analysis/Nut_net_data_mgmt/generated_data/leaf_root.csv", row.names = FALSE)
```



```{r}
se <- function(x) sd(x, na.rm = TRUE) / sqrt(sum(!is.na(x)))

species_mean <- leaf_root %>%
  group_by(Taxon) %>%
  summarise(
    across(c(SRL_cm_g, LDMC_g_g, LA_mm2_no, RTD_gm_mm3, RDMC_g_g, SLA_mm2_mg, RD_cm, height_m),
           list(mean = ~mean(., na.rm = TRUE), se = ~se(.)))
  )

head(species_mean)
```


```{r}
species_mean$Taxon[species_mean$Taxon == "SALSOLA KALI SSP. TRAGUS"] <- "SALSOLA KALI"
species_mean$Taxon[species_mean$Taxon == "AMARANTHUS"] <- "AMARANTHUS SP."
species_mean$Taxon[species_mean$Taxon == "MELILOTUS OFFICINALIS SSP. ALBA"] <- "MELILOTUS OFFICINALIS"
species_mean$Taxon <- trimws(species_mean$Taxon)
```

```{r}
write.csv(species_mean, "C:/My_analysis/Nut_net_data_mgmt/generated_data/trait_mean_se.csv", row.names = FALSE)
```


```{r} 
trait_abun_all <- left_join(abundance_all, species_mean, by = "Taxon")

write.csv(trait_abun_all, "C:/My_analysis/Nut_net_data_mgmt/generated_data/trait_abund_all_2024.csv", row.names = FALSE)
```

```{r}
anti_join(species_mean, abundance_all, by = "Taxon")
```


#find the taxon name that are different that the traits taxon 
```{r}
# Search for each species using partial matching
grep("AMARANTHUS", abundance_all$Taxon, value = TRUE) |> unique()
grep("CAPSELLA", abundance_all$Taxon, value = TRUE) |> unique()
grep("CHONDROSUM", abundance_all$Taxon, value = TRUE) |> unique()
grep("HEMIZONIA", abundance_all$Taxon, value = TRUE) |> unique()
grep("HIERACIUM", abundance_all$Taxon, value = TRUE) |> unique()
grep("LOLIUM", abundance_all$Taxon, value = TRUE) |> unique()
grep("MADIA", abundance_all$Taxon, value = TRUE) |> unique()
grep("MEDICAGO", abundance_all$Taxon, value = TRUE) |> unique()
grep("MELILOTUS", abundance_all$Taxon, value = TRUE) |> unique()
grep("PLANTAGO", abundance_all$Taxon, value = TRUE) |> unique()
grep("SALSOLA", abundance_all$Taxon, value = TRUE) |> unique()
```


